Notes on RBFs methods for Neural Fields

### 2023-11-14 Manuscript/manuscript.tex
First pass at explaining the algorithm

Using \cref{eq:projCollProperty}, we can now pass from \cref{eq:cauchyPn} to an
equivalent ODE in $\RSet^n$. We define the residual
\[
  R \colon C^1(J,\XSet) \to C(J,\XSet), \qquad U \mapsto U' - N(\blank, U),
\]
rewrite \cref{eq:cauchyPn} as
\[
  P_n \bigl( R(U_n)(t) \bigr) = 0, \qquad t \in J, \qquad 
  P_n \bigl( U_n(0) - V\bigr) = 0, 
\]
which, by property \cref{eq:projCollProperty} is equivalent to
\[
  \begin{aligned}
  & U'_n(t)(x_i) = \bigl[P_n N(t,U_n(t))\bigr](x_i), \qquad (i,t) \in \NSet_n \times J, \\
  & U_n(0)(x_i) = (P_n V)(x_i).
  \end{aligned}
\]
Finally, since $U_n(t) \in \XSet_n$, we can write $U_n(t) = \sum_{j \in \NSet_n}
\alpha_j(t) \ell_j$ and obtain the sought ODE on $\RSet^n$,
\begin{equation} \label{eq:NFODE}
  \begin{aligned}
   & \alpha'(t) = -\alpha(t) + K(\alpha(t)) + \gamma(t) \qquad  t \in [0,T],\\ 
   & \alpha (0) = \nu,
  \end{aligned}
\end{equation}  
where 
\[
  \begin{aligned}
  & \alpha_i(t) = U_n(t)(x_i), && \nu_i = v(x_i),  \\
  & \gamma_i(t) = g(x_i,t),
  && K_i(\alpha) = 
    \int_{D}w(x_i,y) f\Bigl( \sum_{j \in \NSet_n} \alpha_j \ell_j(y)\Bigr) \, d\mu(y)
  \end{aligned}
\]
In passing, we note that the system \cref{eq:NFODE} above is related, but not
identical, to \cref{eq:NFODEQuad}. The former is expressed in terms of 
$K \colon \RSet^n \to \RSet^n$, the latter in terms of
$\tilde K \colon \RSet^n \to \RSet^n$, which adopts a quadrature rule to approximate
the integrals in $K$.

\section{RBF scheme for neural field equations}\label{sec:RBFScheme} 
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{sketch}
  \caption{The patch $D_i$ of the spatial domain $D$ contains $k_i =4$}
  \label{fig:sketch}
\end{figure}

\section{Conventions} \label{sec:conventions}
See \texttt{docsiamart.pdf} for many examples. I generally adopt the conventions of
that template. I remark a few things below.

\subsection{References} \label{sec:references} For cross-references I use
\texttt{cleveref}: like this
\cref{eq:test}, like this \Cref{eq:test}, or like this \cref{sec:conventions}, or
this \cref{hyp:trivial}, or this \cref{clm:trivial}.
\begin{equation}\label{eq:test}
  x = y
\end{equation}

\subsection{Todos}
You can find templates of other author's todos in
\texttt{damacros.sty}, which you are welcome to edit.
This is how I make a short comment\da{Short comment by Daniele}. Comments can also go
inline like the one below.
\da[inline]{Long comments by Daniele} 
% \missingfigure
% 
% \section*{Acknowledgments}

\section{Testing commands}

\subsection{Testing maths operators}
Here are tests for operators
\[
  \spec, \quad 
  \range, \quad 
  \spn, \quad
  \diag, \quad
  \meas, \quad
  \id, \quad
  \real z, \quad
  \imag z, \quad
  \e, \quad
  \iunit, \quad
\]

\subsection{Testing additional theorem-like environments} The following additional
theorem-like environments may be useful.

\begin{hypothesis} \label{hyp:trivial}
  This is a hypothesis
\end{hypothesis}

\begin{claim} \label{clm:trivial}
  This is a claim
\end{claim}

\subsection{Testing sets}
This are a few examples of sets (see more commands in \texttt{damacros.sty})
\[
  \RSet, \quad
  \CSet, \quad
  \ZSet, \quad
  \sph, \quad
  \ldots
\]

\subsection{Empty arguments}
Command for empty arguments, which require spacing
\[
  u(\blank,t) \quad t \in \RSet
\]

\subsection{Calligraphic fonts}
If one really must use calligraphic fonts, there are macros for this.
\[
  \calA, \quad
  \calL, \quad
  \calM, \quad
  \calS, \quad
\]

>>>>

Using \cref{eq:projCollProperty}, we can now pass from \cref{eq:cauchyPn} to an
equivalent ODE in $\RSet^n$. We define the residual
\[
  R \colon C^1(J,\XSet) \to C(J,\XSet), \qquad U \mapsto U' - N(\blank, U),
\]
rewrite \cref{eq:cauchyPn} as
\[
  P_n \bigl( R(U_n)(t) \bigr) = 0, \qquad t \in J, \qquad 
  P_n \bigl( U_n(0) - V\bigr) = 0, 
\]
which, by property \cref{eq:projCollProperty} is equivalent to
\[
  \begin{aligned}
  & U'_n(t)(x_i) = \bigl[P_n N(t,U_n(t))\bigr](x_i), \qquad (i,t) \in \NSet_n \times J, \\
  & U_n(0)(x_i) = (P_n V)(x_i).
  \end{aligned}
\]
Finally, since $U_n(t) \in \XSet_n$, we can write $U_n(t) = \sum_{j \in \NSet_n}
\alpha_j(t) \ell_j$ and an ODE on $\RSet^n$,
\begin{equation} \label{eq:NFODE}
  \begin{aligned}
   & \alpha'(t) = -\alpha(t) + K(\alpha(t)) + \gamma(t) \qquad  t \in [0,T],\\ 
   & \alpha (0) = \nu,
  \end{aligned}
\end{equation}  
where 
\[
  \begin{aligned}
  & \alpha_i(t) = U_n(t)(x_i), && \nu_i = v(x_i),  \\
  & \gamma_i(t) = g(x_i,t),
  && K_i(\alpha) = 
    \int_{D}w(x_i,y) f\Bigl( \sum_{j \in \NSet_n} \alpha_j \ell_j(y)\Bigr) \, d\mu(y)
  \end{aligned}
\]
In passing, we note that the system \cref{eq:NFODE} above is related, but not
identical, to \cref{eq:NFODEQuad}. The former is expressed in terms of 
$K \colon \RSet^n \to \RSet^n$, the latter in terms of
$\tilde K \colon \RSet^n \to \RSet^n$, which adopts a quadrature rule to approximate
the integrals in $K$. In the next section, we will give a concrete example of $\tilde
K$

\section{RBF scheme for neural field equations}\label{sec:RBFScheme} 
\begin{figure}
  \centering
  \includegraphics[width=\textwidth]{sketch}
  \caption{The patch $D_i$ of the spatial domain $D$ contains $k_i =4$}
  \label{fig:sketch}
  \da[inline]{We need a meaningful sketch here}
\end{figure}

We now introduce the interpolant we will use in the paper. To define the interpolant
we start from the set of $n$ nodes $\Xi = \{ x_i \}_{i \in \NSet_n}$, with the
associated \textit{Voronoi diagram} $\{ D_i \}_{i \in \NSet_n}$, that is, the
partition of $D$ into cells
\[
  D_i = \Bigl\{ x \in D \colon \| x - x_i \| = \min_{j \in \NSet_n} \| x - x_j \| \Bigr\}.
\]
Further, we consider a map associating each node $x_i$ to $k_i$ nodes $\{\xi_{ij}
\}_{j = 1}^{k_i}$ in $\Xi$, the stencil of $x_i$. More formally we define the
mapping
\[
  S \colon \Xi \to 2^{\,\Xi}, \qquad x_i \mapsto \{ \xi_{ij} \}_{j = 1}^{k_i},
\]
where $2^{\,\Xi}$ is the power set $\Xi$, and denote by $S(x_i)$, or simply $S_i$, the
stencil of node $i$. We stipulate that $x_i$ is part of its stencil, $x_i \in
S(x_i)$. Finally, we define a basic radial function $\phi = \Phi( \| \blank \|)
\colon \RSet^d \to \RSet$, and construct an interpolant of $v$ in the
following form
\begin{equation}\label{eq:RBFExpStencil}   
  v_n(x) = \sum_{j = 1}^{n} \sum_{l=1}^{k_j} V_{jl} \Phi(\| x - \xi_{jl} \|) 1_{D_j}(x) 
\end{equation}
where $1_A$ denotes the indicator function of the set $A \subset \RSet^d$. In
\cref{eq:RBFExpStencil}, the interpolant sums over the stencil in each node, and it
is therefore expressed in terms of $k_1 + \dots + k_n \geq n$ coefficients $\{
V_{jl}\}$. Since $v_n$ interpolates $v$ at the nodes $\Xi$, the coefficients
$V_{jl}$ satisfy the $n$ equations
\[
  v(x_j) = \sum_{l=1}^{k_j} V_{jl} \Phi(\| x_j - \xi_{jl} \|), \qquad j \in \NSet_n.
\]
The system above does not uniquely specify the coefficients $V_{jl}$, and so further
constraints are added,
\da[inline]{Complete this after Sage's input}

Crucially, it is possible to pass from the form \cref{eq:RBFExpStencil} of the
interpolant, to the so-called \textit{cardinal form} of the interpolant, that, is a
Lagrange interpolant
\[
  v_n(x) = \sum_{j=1}^{n} v(x_j) \ell_j(x),
\]
using the transformation
\da[inline]{Complete after Sage's input}

Henceforth, we will use the cardinal form for the analysis and the presentation of
our scheme, while we resort to the stencil-based form to carry out concrete
calculations. The cardinal form of the interpolant defines a projector $P_n \colon \XSet \to
\XSet_n = \spn \{ \ell_1, \ldots, \ell_n \}$, and therefore fits into the framework
presented in the past section.

To complete the scheme, we now require a quadrature formula for the integrand. The
formula we employ is a composite formula over the Voronoi diagram
\[
  \begin{aligned}
  Q v := \int_{D} v(y)\,d \mu(y) 
    & = \sum_{j=1}^n \int_{D_j} v(y) \,d \mu(y) \\
    & \approx \sum_{j = 1}^n \int_{D_j} v_n(y) \,d \mu(y)
      \approx \sum_{j = 1}^n v(x_j) \mu_j =: Q_n v,
  \end{aligned}
\]
where the weights $\{ \mu_j \}_{j \in \NSet_n}$ satisfy
\da[inline]{Complete. I am unsure there are two approximation symbols above, or just
one. If possible, say something about the expected convergence rate of $|Qv - Q_nv|$.} 
We are now ready to specify a discrete scheme by applying the quadrature rule $Q_n$
the function $K$ in \cref{eq:NFODE}, recalling that each $D_i$ contains exactly
one node, namely $x_i$, and obtaining the spatially-discrete scheme
\[
  \begin{aligned}
  & \tilde \alpha'(t) = -\tilde \alpha(t) + \gamma(t) 
    + M f(\tilde \alpha(t)), \qquad t \in J \\
  & \tilde \alpha(0) = \nu
  \end{aligned}
\]
where $M_{ij} = w(x_i,x_j)\mu_j$, and in which we have used, with a little abuse of
notation, $f(\alpha)$ to denote the vector with components $(f(\alpha))_i =
f(\alpha_i)$. We have therefore provided an RBF example of the collocation scheme
\cref{eq:NFODEQuad}, in which $\tilde K(\tilde \alpha) = M f(\tilde \alpha)$.
